---
title: "Lab3B"
author: "Simin Manole"
date: "September 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

## Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight.
```{r}
surveys <- read_csv("data/portal_data_joined.csv")
surveys_ysw <- surveys %>%
  filter(year < 1995) %>%
  select(year, sex, weight)

surveys_ysw
```

## Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3.
```{r}
surveys_hindfoot <- surveys %>%
  mutate(hindfoot_cm = hindfoot_length / 10) %>%
  filter(!is.na(hindfoot_cm)) %>%
  filter(hindfoot_cm < 3) %>%
  select(species_id, hindfoot_cm)
```

## How many animals were caught in each plot_type surveyed?
```{r}
surveys %>%
  count(plot_type)
```

## Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).
```{r}
surveys %>%
  group_by(species_id) %>%
  filter(!is.na(hindfoot_length)) %>%
  summarize(mean_hindfoot = mean(hindfoot_length),
            min_hindfoot = min(hindfoot_length),
            max_hindfoot = max(hindfoot_length),
            n = n()
            )

```

## What was the heaviest animal measured in each year?
```{r}
surveys %>%
  group_by(year) %>%
  filter(!is.na(weight)) %>%
  filter(weight == max(weight)) %>%
  select(year, genus, species, weight) %>%
  arrange(year)
  
```

## Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values.
```{r}
surveys_spread_genera <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(genera = n_distinct(genus)) %>%
  #Year is key
  #number of genera is value
  spread(key = year, value = genera)

  head(surveys_spread_genera)
```

## Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.
```{r}
surveys_spread_genera %>%
  #year is key
  #number of genera is value
  gather(year, genera, -plot_id)
```

## Use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight.

First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. 
```{r}
surveys_measurements <- surveys %>%
  gather(hindfoot_length, weight, key = measurement, value = value)
```

## With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. 
Hint: You only need to specify the key and value columns for spread().
```{r}
surveys_measurements_mean <- surveys_measurements %>%
  filter(!is.na(value)) %>%
  select(year, plot_type, measurement, value) %>%
  group_by(plot_type, year, measurement) %>%
  summarize(mean = mean(value)) %>%
  spread(key = measurement, value = mean)
  head(surveys_measurements_mean)
```