---
title: "Lab3B"
author: "Simin Manole"
date: "September 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

## Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight.
```{r}
surveys <- read_csv("data/portal_data_joined.csv")
surveys_ysw <- surveys %>%
  filter(year < 1995) %>%
  select(year, sex, weight)

surveys_ysw
```

## Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3.
```{r}
surveys_hindfoot <- surveys %>%
  mutate(hindfoot_cm = hindfoot_length / 10) %>%
  filter(!is.na(hindfoot_cm)) %>%
  filter(hindfoot_cm < 3) %>%
  select(species_id, hindfoot_cm)
```

## How many animals were caught in each plot_type surveyed?
```{r}
surveys %>%
  count(plot_type)
```

## Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).
```{r}
surveys %>%
  group_by(species_id) %>%
  filter(!is.na(hindfoot_length)) %>%
  summarize(mean_hindfoot = mean(hindfoot_length),
            min_hindfoot = min(hindfoot_length),
            max_hindfoot = max(hindfoot_length),
            n = n()
            )

```

## What was the heaviest animal measured in each year?
```{r}
surveys %>%
  group_by(year) %>%
  filter(!is.na(weight)) %>%
  filter(weight == max(weight)) %>%
  select(year, genus, species, weight) %>%
  arrange(year)
  
```

## Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values.

You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.

plot_id is key
number of genera is value

need to make genus to value
```{r}
surveys %>%
  #select(year, plot_id, genus)
  count(plot_id, genus)
```