---
title: "Lab3A"
author: "Simin Manole"
date: "September 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, eval=FALSE}
download.file(url="https://ndownloader.figshare.com/files/2292169/combined.csv",
              destfile = "data/portal_data_joined.csv")
```

```{r}
surveys <- read.csv("data/portal_data_joined.csv")
```

## What is the class of the object surveys?
```{r}
str(surveys)
##class = 'data.frame'
```
## How many rows and how many columns are in this object?
```{r}
ncol(surveys)
nrow(surveys)
##number of columns = 13
##number of rows = 34786
```
## How many species have been recorded during these surveys?
```{r}
length(summary(surveys[ , "species_id"]))
##number of species = 48
```

## Create a data.frame containing containing only the data in row 200 of the surveys dataset.
```{r}
surveys_200 <- surveys[200, ]
```

## Create a new data frame (surveys_last) from that last row.
```{r}
nrow(surveys)
tail(surveys)
n_rows <- nrow(surveys)
surveys_last <- surveys[n_rows,]
```

## Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.
```{r}
surveys_middle <- surveys[n_rows /2 ,]
```

## Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.
```{r}
surveys_head <- surveys[-(7:n_rows),]
```

## Rename “F” and “M” to “female” and “male” respectively.
```{r}
sex <- factor(surveys$sex)
levels(sex)[1] <- "undetermined"
levels(sex)[2] <- "female"
levels(sex)[3] <- "male"
levels(sex)
```

## Can you recreate the barplot such that “undetermined” is last (after “male”)?
```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```

## There are a few mistakes in this hand-crafted data.frame. Can you spot and fix them? Don’t hesitate to experiment!
```{r, eval=FALSE}
animal_data <- data.frame(
          animal = c(dog, cat, sea cucumber, sea urchin),
          feel = c("furry", "squishy", "spiny"),
          weight = c(45, 8 1.1, 0.8)
          )
```

```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "furry", "squishy", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
          )
```

## Can you predict the class for each of the columns in the following example?
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
str(country_climate)
```

```{r}
country_climate <- data.frame(
  country = c("Canada", "Panama", "South Africa", "Australia"),
  climate = c("cold", "hot", "temperate", "hot/temperate"),
  temperature = c(10, 30, 18, 15),
  northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
  has_kangaroo = c(FALSE, FALSE, FALSE, TRUE),
  stringsAsFactors = FALSE
  )
str(country_climate)
```
