---
title: "Lab4 ggplot2 Exercises"
author: "Simin Manole"
date: "September 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("DT")
library("hexbin")
surveys_complete <- read_csv("data/portal_data_joined.csv")
```

## Example Plot
```{r}
# Assign plot to a variable
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

# Draw the plot
surveys_plot + 
    geom_point()
```

## Use the geom_hex() function
```{r}
surveys_plot +
 geom_hex()
```
```{r}
#geom_hex adds a new dimension so I can see areas of high overlap, but I lose resolution in my x & y points, so say the grouping above hindfoot_length 40 is less compact and now attaches to other groups.
```

## Use what you just learned to create a scatter plot of weight over species_id with the plot types showing in different colors
```{r}
#Added to remove NA data from plots
surveys_ggplot <- surveys_complete %>%
  filter(!is.na(weight))

ggplot(data = surveys_ggplot, 
        mapping = aes(x = species_id, y = weight)) +
  geom_point(aes(color = plot_type))
```
Is this a good way to show this type of data?
```{r}
#What a mess! You're better off doing average weight per plot type to clean up this mix of colored dots, and still I'd say a bar chart makes more sense.
```

## Replace the box plot with a violin plot
```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
    geom_violin(alpha = 0)
```

## Represent weight in log10 scale
```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
  geom_violin(alpha = 0) +
  scale_y_log10()
```

## Create a boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements.
```{r}
surveys_ggplot_2 <- surveys_complete %>%
  filter(!is.na(hindfoot_length))

ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
   geom_boxplot(alpha = 0)
```

## Add color to the data points on your boxplot according to the plot from which the sample was taken (plot_id).
```{r}
ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```

## But if plot_id was a factor?
```{r}
#set as factor 
surveys_ggplot_2 <- mutate_at(surveys_ggplot_2, vars(plot_id), as.factor)
#plot
ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```




