---
title: "Lab 4"
author: "Simin Manole"
date: "September 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("DT")
```

## Import the JHU case tracking data 
```{r}
report_03_11_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

## Interactive Data Table
```{r}
datatable(report_03_11_2020)
```

## What are the differences between 3/13/2020, the Friday before spring break 2020 and yesterday 6 months later on 9/13/2020? 
```{r}
report_09_13_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv"))
str(report_03_11_2020)
str(report_09_13_2020)
```

In the six months between the two reports the data now includes new columns such as FIPS, Admin2, Active, Incidence_Rate, Combined_Key, and 'Case-Fatality_Ratio'.

## Make a data frame (tibble) with total counts for each country.
```{r}
report_March_country_total <- report_03_11_2020 %>%
  group_by(Country_Region) %>%
  summarise(
    Total_Confirmed = sum(Confirmed),
    Total_Deaths = sum(Deaths),
    Total_Recovered = sum(Recovered)
  )
report_Sept_country_total <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarise(
    Total_Confirmed = sum(Confirmed),
    Total_Deaths = sum(Deaths),
    Total_Recovered = sum(Recovered)
  )
datatable(report_March_country_total)
datatable(report_Sept_country_total)
```

## Now just the USA for confirmed cases per each state
```{r}
report_March_US_conf <- report_03_11_2020 %>%
  filter(Country_Region == "US") %>%
  select(Province_State, Confirmed)

report_Sept_US_conf <- report_09_13_2020 %>%
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>%
  summarize(Total_Confirmed = sum(Confirmed))

datatable(report_March_US_conf)
datatable(report_Sept_US_conf)
```

## Let's just use the top 10 countries for Deaths
```{r}
report_March_top10 <- report_March_country_total %>%
  select(Country_Region, Total_Deaths) %>%
  arrange(desc(Total_Deaths)) %>%
  slice(1:10)
report_Sept_top10 <- report_Sept_country_total %>%
  select(Country_Region, Total_Deaths) %>%
  arrange(desc(Total_Deaths)) %>%
  slice(1:10)
datatable(report_March_top10)
datatable(report_Sept_top10)
```

## Make plots using geom_point() of the (top) 10 countries with the confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.
```{r}
#Let's do deaths first since we have that data ready
March_death_plot <- ggplot(data = report_March_top10, 
                       mapping = aes(x = Country_Region, y = Total_Deaths))
March_death_plot + 
    geom_point() +
  ggtitle("Total Deaths per Country, March 2020")

Sept_death_plot <- ggplot(data = report_Sept_top10, 
                       mapping = aes(x = Country_Region, y = Total_Deaths))
Sept_death_plot + 
    geom_point() +
  ggtitle("Total Deaths per Country, September 2020")
#Don't know how to reorder X-axis labels so Yis in descending order

#Now confirmed cases
report_March_top10_conf <- report_March_country_total %>%
  select(Country_Region, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)
report_Sept_top10_conf <- report_Sept_country_total %>%
  select(Country_Region, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)

#Now graph it
March_conf_plot <- ggplot(data = report_March_top10_conf, 
                       mapping = aes(x = Country_Region, y = Total_Confirmed))
March_conf_plot + 
    geom_point() +
  ggtitle("Top 10 Total Confirmed Cases per Country, March 2020")

Sept_conf_plot <- ggplot(data = report_Sept_top10_conf, 
                       mapping = aes(x = Country_Region, y = Total_Confirmed))
Sept_conf_plot + 
    geom_point() +
  ggtitle("Top 10 Total Confirmed Cases per Country, September 2020")
```
 
## Make using geom_bar of the (top) 10 states with the most confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.
```{r}
report_March_top_US_conf <- report_March_US_conf %>%
  select(Province_State, Confirmed) %>%
  arrange(desc(Confirmed)) %>%
  slice(1:10)

ggplot(report_March_top_US_conf, aes(x = Confirmed, y = Province_State)) +
geom_bar(stat = 'identity') +
  ggtitle("Top 10 Total Confirmed Cases per US State, March 2020")

report_Sept_top_US_conf <- report_Sept_US_conf %>%
  select(Province_State, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)

ggplot(report_Sept_top_US_conf, aes(x = Total_Confirmed, y = Province_State)) +
geom_bar(stat = 'identity') +
  ggtitle("Top 10 Total Confirmed Cases per US State, September 2020")
```


