---
title: "Lab 5"
author: "Simin Manole"
date: "September 21, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("DT")
surveys_complete <- read.csv("data/portal_data_joined.csv")
```


## Replace the box plot with a violin plot; see geom_violin()
```{r}
#Need to remove NA to look decent
surveys_weight <- surveys_complete %>% filter(!is.na(weight))
#Then plot
ggplot(data = surveys_weight, mapping = aes(x = species_id, y = weight)) +
    geom_violin()
```

## Represent weight on the log10 scale
```{r}
ggplot(data = surveys_weight, mapping = aes(x = species_id, y = weight)) +
    geom_violin() +
  scale_y_log10()
```

## Create a boxplot for hindfoot_length, then overlay the boxplot layer on a jitter layer to show actual measurements. Finally, add color to the data points on your boxplot according to the plot from which the sample was taken (plot_id).
```{r}
surveys_ggplot_2 <- surveys_complete %>%
  filter(!is.na(hindfoot_length))

ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
   geom_boxplot(alpha = 0)#set as factor 
```

```{r}
surveys_ggplot_2 <- mutate_at(surveys_ggplot_2, vars(plot_id), as.factor)
ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```

## Use what you just learned to create a plot that depicts how the average weight of each species changes through the years.
```{r}
yearly_weight <- surveys_weight %>%
  group_by(year, species_id) %>%
  summarize(mean_weight = mean(weight))
```
```{r}
ggplot(data = yearly_weight, aes(x = year, y = mean_weight)) +
     geom_line() +
  facet_wrap(facets = vars(species_id)) +
  theme_bw()
```

##  Improve one of the plots generated in this exercise or create a beautiful graph of your own.
```{r}
#Let's fix the last graph
ggplot(data = yearly_weight, aes(x = year, y = mean_weight)) +
  geom_line() +
  facet_wrap(facets = vars(species_id)) +
  labs(title = "Observed Weight Through Time by Species",
        x = "Year of observation",
        y = "Average Weight") +
  theme(axis.text.x = element_text(
        colour = "grey20", size = 10, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        strip.text = element_text(face = "italic"),
        text = element_text(size = 12))
```

## Revise the above example for joining tables using 6/13/2020 and 9/13/2020. 
```{r, message = FALSE}
#Let's get the tables ready
Confirmed_State_6_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/06-13-2020.csv")) %>%
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed))
#Small fix
Confirmed_State_6_13 <- Confirmed_State_6_13 %>% 
  filter(Province_State != "Recovered") 

Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```
```{r}
#Now to join them
Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_6_13_9_13_joined)
```
```{r}
#Fix those columns
Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_6_13_2020 = 0))
head(Confirmed_State_6_13_9_13_joined)
```
```{r}
#Now to make the table 'long'
Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```

## Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height in the R code chunk header.
```{r, fig.width = 10, fig.height = 10}
ggplot(data = Confirmed_State_6_13_9_13_joined_long, aes(y = Province_State, x = Confirmed)) +
    labs(title = "Confirmed Cases within The USA",
        y = "State",
        x = "Confirmed Cases") +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 12),
    strip.text = element_text(face = "italic"),
    text = element_text(size = 20)) +
  geom_bar(stat='identity', aes(color = Date))
```

## Using the time series data, make a plot of the total number of confirmed cases & deaths, per day, worldwide.
```{r, message = FALSE}
#Get started with example data
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", destfile = "data/time_series_covid19_confirmed_global.csv")

time_series_confirmed <- read_csv("data/time_series_covid19_confirmed_global.csv") %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
#Make it long!
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 
```

```{r}
#Now the same but with deaths
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", destfile = "data/time_series_covid19_deaths_global.csv")

time_series_deaths <- read_csv("data/time_series_covid19_deaths_global.csv") %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 

head(time_series_deaths_long)
```

```{r}
#Let's combine confirmed & deaths
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)
```
```{r}
#Quick formating stuff
library(lubridate)
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```

```{r}
#Let's finally make that plot
#Make a plot of the total number of confirmed cases & deaths, per day, worldwide
time_series_long_joined_counts %>% 
  group_by(Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
    ggplot(aes(x = Date,  y = Counts, fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("Worldwide COVID-19 Cases")
```

## Use Mutate to make a new column with deaths/confirmed cases
```{r}
time_series_long_joined_ratio <- time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise(Total_Confirmed = sum(Confirmed), Total_Deaths = sum(Deaths)) %>%
  mutate(Ratio = Total_Deaths / Total_Confirmed)
```
## Plot US deaths/confirmed cases per day.
```{r}
time_series_long_joined_ratio %>%
    filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Ratio)) + 
    geom_point() +
    geom_line() +
    ggtitle("Ratio of Deaths to Confirmed Cases in the US")
```

## Make a single graph with the times series data with the 10 countries with the hightest death total.
```{r}
#Find highest deaths first
time_series_long_total_deaths <- time_series_long_joined_ratio %>%
  group_by(Country_Region) %>%
  summarise(Deaths = sum(Total_Deaths)) %>%
  arrange(desc(Deaths)) %>%
  slice(1:10)
datatable(time_series_long_total_deaths)

#Now we can manually select theses countries for the plot
time_series_long_joined_ratio %>% 
    filter(Country_Region %in% c("US", "Brazil", "United Kingdom", "Italy", "Mexico", "France", "Spain", "India", "Iran", "Peru")) %>% 
    ggplot(aes(x = Date,  y = Total_Deaths, fill = Country_Region, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("The Top 10 Countries by Total Deaths")

```
## Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total.
```{r, fig.height = 8}
#```{r, fig.height = 8}
time_series_long_joined_ratio %>% 
    filter(Country_Region %in% c("US", "Brazil", "United Kingdom", "Italy", "Mexico", "France", "Spain", "India", "Iran", "Peru")) %>% 
    ggplot(aes(x = Date,  y = Total_Deaths, fill = Country_Region, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("The Top 10 Countries by Total Deaths") +
  facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

## Now solved another way thanks to Piazza!
```{r, fig.height = 8}
ten_highest_deaths_df<-time_series_long_joined %>% 
  group_by(Country_Region) %>% 
  summarise("Death_Total"= sum(Deaths)) %>% 
  arrange(desc(Death_Total)) %>% 
  head(10) 

time_series_long_joined_ratio %>% 
  filter(Country_Region %in% ten_highest_deaths_df$Country_Region) %>% 
    ggplot(aes(x = Date,  y = Total_Deaths, fill = Country_Region, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("The Top 10 Countries by Total Deaths") +
  facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

