---
title: "Lab 9x"
author: "Simin Manole"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Conditional statements with if-else
A basic "if" statement if condition is true
```{r, eval=FALSE}
if(<condition>) {
        ## do something
} 
## Continue with rest of code
```

An "if" statement with a function for condition being false
```{r, eval=FALSE}
if(<condition>) {
        ## do something
} 
else {
        ## do something else
}
```

Testing two "if" statements in series with a function for both being false.
```{r, eval=FALSE}
if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}
```

An example:
  "Your character is unconscious. Roll a d20 so see if they wake up or not. Roll a ten or higher to make the save."
```{r}
## Generate a uniform random number
x <- runif(1, 1, 20)  
if(x > 9.999) {
        y <- "Save"
} else {
        y <- "Fail"
}

if(y == "Save") {
        z <- "You survive."
} else {
        z <- "Yeah that's not happening buddy."
}
```

"Your roll?"
```{r}
x
```

"That's a..." 
```{r}
y
```

"Did I make it?"
```{r}
z
```

## "for" loops
These take an iterating variable and assign it to successive values from a sequence or vector. Here i is given a value from the sequence 1:10 each loop and is printed.
```{r}
for(i in 1:5) {
        print(i)
}
```

This loop does the same, but the sequence is created out of the loop, then called within it.
```{r}
x <- c("z", "y", "x", "w")
for(i in 1:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

This loop is similar to the previous but the length of iteration is set to the exact lenght of the sequence
```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```

Alternatively, search by type and not an index-type variable.
```{r}
for(letter in x) {
        print(letter)
}
```

One line loops don't require the curly braces.
```{r}
for(i in 1:4) print(x[i])
```

## Nested "for" loops
Nested loops are commonly used for multidimensional or hierarchical data structures. Here, a 2x3 matrix is made, and one loop calls the row number then the next prints data in each column, then 
```{r}
x <- matrix(1:9, 3, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```

## "while" Loops

Here the loop will loop until a condition is true. This example increases the value of "count" each loop and exits once it reaches a value not less than 10.
```{r}
count <- 0
while(count < 5) {
        print(count)
        count <- count + 1
}
```

Here there are two conditions in the test. The order is important, as here "z>=3" must be TRUE before the next condition is tested.
```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  

## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```

## Your First Function
Here’s a simple function that does nothing. We create a function "f" that gives no result.
```{r, eval=FALSE}
f <- function() {
## This is an empty function
}

## Functions have their own class
class(f)  

## Execute this function
f()       
```

The next thing we can do is create a function that actually has a non-trivial function body. Aka is does something.
```{r,}
f <- function() {
         cat("I'm sorry, Dave. I'm afraid I can't do that.\n")
}
f()
```

Adding a function argument allows a more flexible function. Instead of calling a function multiple times, here a numerical argument will loop function f as many times as the argument that is input.
```{r}
f <- function(num) {
        for(i in seq_len(num)) {
                cat("I'm sorry, Dave. I'm afraid I can't do that.\n")
        }
}
f(3)
```

For the following example, function f the characters of "Hello, world!" are repeated according to the numerical argument and the last thing called is "chars" that gives that value. f(3) then will count the string 3 times and give the number of characters.
```{r}
f <- function(num) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(3)
```

```{r}
print(meaningoflife)
```

Since our function looks for an argument a simple f() will cause an error. We can create a default value that is used in the abscesence of a user defined value.
```{r}
f <- function(num = 1) {
        sorry <- "I'm sorry, Dave. I'm afraid I can't do that.\n"
        for(i in seq_len(num)) {
                cat(sorry)
        }
        chars <- nchar(sorry) * num
        chars
}
f()    ## Use default value for 'num'
```

But a User defined value has priority.
```{r}
f(2)
```

Alternatively, we can specify the argument "num" from the funcion specifically.
```{r}
f(num = 2)
```

