---
title: "Lab 9x"
author: "Simin Manole"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Conditional statements with if-else
A basic "if" statement if condition is true
```{r, eval=FALSE}
if(<condition>) {
        ## do something
} 
## Continue with rest of code
```

An "if" statement with a function for condition being false
```{r, eval=FALSE}
if(<condition>) {
        ## do something
} 
else {
        ## do something else
}
```

Testing two "if" statements in series with a function for both being false.
```{r, eval=FALSE}
if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}
```

An example:
  "Your character is unconscious. Roll a d20 so see if they wake up or not. Roll a ten or higher to make the save."
```{r}
## Generate a uniform random number
x <- runif(1, 1, 20)  
if(x > 9.999) {
        y <- "Save"
} else {
        y <- "Fail"
}

if(y == "Save") {
        z <- "You live."
} else {
        z <- "Yeah that's not happening buddy."
}
```

"Your roll?"
```{r}
x
```

"That's a..." 
```{r}
y
```

"Did I make it?"
```{r}
z
```

## Your First Function
Here’s a simple function that takes no arguments and does nothing.
```{r, eval=FALSE}
f <- function() {
       ## This is an empty function
}

## Functions have their own class
class(f)  
[1] "function"

## Execute this function
f()       
NULL
```

The next thing we can do is create a function that actually has a non-trivial function body.
```{r,}
f <- function() {
         cat("Hello, world!\n")
}
f()
```